---
title: "WDFW Budget"
author: "Ian Bogley"
date: "12/8/2021"
output: pdf_document
---

Today we are going to set out to visualize budget data for the Washington Department of Fish and Wildlife.

I will be getting my data from their page on the 2022 legislative session: https://wdfw.wa.gov/about/administration/budget/update

To start, let's load in our packages.

```{r intro}
library(pacman)
p_load(tidyverse,ggrepel)
```

In the future, perhaps I will use Rvest and httr to webscrape this information, but for now let's manually input the data.

```{r data}
#WDFW 2022 operating funding emergency needs budget
df_operating_funding_emergency_needs <- data.frame(project=c("Supporting habitat-friendly solar facilities",
                                                             "Forage fish spawning monitoring",
                                                             "Integrating salmon recovery plans at local level",
                                                             "Freshwater monitoring",
                                                             "Meeting recreation needs",
                                                             "Building salmon team capacity",
                                                             "Fish passage rulemaking",
                                                             "License reduction and alternative gear: Columbia River",
                                                             "Hatchery production and compliance",
                                                             "Monitoring recreational shellfish harvest",
                                                             "Safe and sanitary water access areas",
                                                             "Improving wildlife monitoring",
                                                             "Supplemental maintenance level packages"),
                                                   funds = c(400000,700000,
                                                             1300000,2600000,
                                                             3500000,1000000,
                                                             300000,3400000,
                                                             600000,300000,
                                                             500000,2400000,
                                                             24000000))

#WDFW 2022 Capital projects budget
df_capital_projects <- data.frame(
  projects = c("Duckabush Estuary Restoration Project",
               "Hoffstadt Hills Aqcuisition",
               "Deschutes Watershed Center Hatchery Work"),
  funds = c(51000000,14500000,2200000))
```

Now let's visualize the operating funding emergency budget in a pie chart.

```{r operating}
#Plotting operating emergency funding needs
p1 <- df_operating_funding_emergency_needs %>% 
  ggplot(aes(x = "", y = funds, fill = project)) + 
  geom_bar(stat = "identity",width = 1,color = "black") +
  coord_polar("y",start = 0) +
  theme_void() + 
  labs(title = "2022 Legislative Session: WDFW", 
       subtitle = paste("Operating Funding Emergency Needs, Total = $",
                          format(sum(df_operating_funding_emergency_needs$funds),
                                 scientific = FALSE,big.mark = ","),sep = ""),
       fill = "") +
  scale_fill_brewer(palette = "Set3")

p1
```

